A performance comparison of Dask and Apache Spark for data-intensive neuroimaging pipelines

### 2019-09-27 latex/paper.tex
Added paragraph on Impact of the NFS of the different applications

\author{Mathieu Dugr\'e, Val\'erie Hayot-Sasson, Tristan Glatard\\
Department of Computer Science and Software Engineering\\
Concordia University, Montr\'eal, Qu\'ebec, Canada 
\vspace*{0.8cm} % to avoid weird spacing of 1st page by Latex.
}

@@

First and foremost, Spark is written in Scala while Dask is in Python.
Given the popularity of Python in scientific communities, this arguably
gives an edge to Dask due to data serialization costs from Python to Scala.
On the other hand, Python's Global Interpreter Lock (GIL) might reduce
parallelism in some cases. This difference in programming languages also
has qualitative implications. As part of the \href{http://scipy.org}{SciPy}
ecosystem, Dask provides almost transparent parallelization of applications
manipulating Numpy arrays or Pandas data frames. On the other hand, Spark's
Java, R and Python APIs allow to easily parallelize analyses that combine
these languages, with reduced performance loss. Our study focuses on
performance, although we recognize that other factors will also impact the
choice of an engine in practice.

@@

@@

Although Dask reduces the impact of Python's GIL on parallelism by using external
Python extensions, our Python histogram experiment (Figure
~\ref{fig:histo_worker} and \ref{fig:histo_block}) showed that issues remain in Dask to
parallelize Python functions. However, Dask was easily able to parallelize
the same application when implemented with Numpy
(Figure~\ref{fig:histo_np_worker} and \ref{fig:histo_np_block}) because in this case the computation
happened in C, outside of the GIL. Similarly, the computation in
command-line applications happens in a sub-process, hence outside of the
GIL, which explains why the BIDS apps example application was not impacted
by this issue. Although implemented in Python, Incrementation was not
impacted either, as computation was only emulated through sleep time.

>>>>

\author{Mathieu Dugr\'e, Val\'erie Hayot-Sasson, Tristan Glatard\\
Department of Computer Science and Software Engineering\\
Concordia University, Montr\'eal, Qu\'ebec, Canada\\
\{mathieu.dugre, valerie.hayot-sasson, tristan.glatard\}@concordia.ca
\vspace*{0.8cm} % to avoid weird spacing of 1st page by Latex.
}

@@

First and foremost, Spark is written in Scala while Dask is in Python.
Given the popularity of Python in scientific communities, this arguably
gives an edge to Dask due to data serialization costs from Python to Scala.
On the other hand, Python's Global Interpreter Lock (GIL) might reduce
parallelism in some cases. This difference in programming languages also
has qualitative implications. As part of the \href{http://scipy.org}{SciPy}
ecosystem, Dask provides almost transparent parallelization of applications
manipulating NumPy arrays or Pandas data frames. On the other hand, Spark's
Java, R and Python APIs allow to easily parallelize analyses that combine
these languages, with reduced performance loss. Our study focuses on
performance, although we recognize that other factors will also impact the
choice of an engine in practice.

@@

Throughout the applications, the use of large data blocks amplified the consequences
of using an NFS. Moreover, the bandwidth saturation almost neglects the scaling, from
additional workers, for data-intensive applications: Incrementation and Histogram
(NumPy). On the other hand, an almost linear scaling shows that it has no substantial
effect on compute-intensive applications: Histogram (Python) and BIDS app example.

We suppose that using a parallel file system would allow applications to have almost
linear scaling as long as they do not saturate the bandwidth of the system.

@@

Although Dask reduces the impact of Python's GIL on parallelism by using external
Python extensions, our Python histogram experiment (Figure
~\ref{fig:histo_worker} and \ref{fig:histo_block}) showed that issues remain in Dask to
parallelize Python functions. However, Dask was easily able to parallelize
the same application when implemented with NumPy
(Figure~\ref{fig:histo_np_worker} and \ref{fig:histo_np_block}) because in this case the computation
happened in C, outside of the GIL. Similarly, the computation in
command-line applications happens in a sub-process, hence outside of the
GIL, which explains why the BIDS apps example application was not impacted
by this issue. Although implemented in Python, Incrementation was not
impacted either, as computation was only emulated through sleep time.
